---
title: "project1"
author: "Safwan"
date: "02/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#echo=TRUE is set as default
library("knitr")

```

## R Markdown
#steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
#2. date: The date on which the measurement was taken in YYYY-MM-DD format
#3. interval: Identifier for the 5-minute interval in which measurement was taken

#Loading and preprocessing data
```{r processing}

data <- read.csv("activity.csv",header = T)
dim(data)
head(data)

```
#Total number of steps taken per day
```{r total steps}
library(ggplot2)
library(plyr)
totalNumberofSteps <- with(data, tapply(steps, as.factor(data$date), sum, na.rm = T))
hist(totalNumberofSteps, main = "HISTOGRAM OF TOTAL NUMBER OF STEPS PER DAY", xlab = "TOTAL NUMBER OF STEPS")
summary(totalNumberofSteps) #to get mean and median
```
#average daily pattern
```{r stepspertime}
library(dplyr)
StepsPerTime <- aggregate(steps~interval,data=data,FUN=mean,na.action=na.omit)
StepsPerTime$time <- StepsPerTime$interval/100
# draw the line plot
h <- ggplot(StepsPerTime, aes(time, steps))
h+geom_line(col="brown")+ggtitle("AVERAGE STEPS PER TIME INTERVAL")+xlab("TIME")+ylab("STEPS")+theme(plot.title = element_text(face="bold", size=14))
# table for dplyr
ST <- tbl_df(StepsPerTime)
# find the column
ST %>% select(time, steps) %>% filter(steps==max(ST$steps))


```
#Imputing missing values
```{r missing values}

MissingData <- data[is.na(data$steps),]
dim(MissingData )


length(MissingData$steps)
MeanSteps <- with(data, tapply(steps, data$interval, mean))
MissingData$steps <-MeanSteps
NewData <- rbind(data,MissingData)
NewData <- NewData[order(NewData$date), ]
TotalNumberofSteps <- with(NewData, tapply(steps, as.factor(NewData$date), sum))
#Make a histogram of the total number of steps taken each day
hist(TotalNumberofSteps, main = "HISTOGRAM OF TOTAL NUMBER OF STEPS PER DAY", xlab = "TOTAL NUMBER OF STEPS")
summary(TotalNumberofSteps)

```
#Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
NewData$days <- weekdays(as.Date(NewData$date))
# find weekend features in the dataset
Weekend_Feature <- grep("Saturday|Sunday", NewData$days, ignore.case = T)
# subset data of the weekend
Weekend_Data<-  NewData[Weekend_Feature, ]
Weekend_Data$weekday <- "weekend"

# subset data of the weekday
Weekday_Data <- subset(NewData,NewData$days!=Weekend_Feature)

Weekday_Data$weekday <- "weekday"

NewData2 <- rbind(Weekday_Data, Weekend_Data)
mean_number_steps <- aggregate(steps~ interval+weekday, NewData2, mean)
g <- qplot(interval, steps, data = mean_number_steps, facets = weekday~.)
g + geom_line(size = 1) + ylab("MEAN STEPS") + ggtitle("AVERAGE NUMBER OF STEPS TAKEN \n AVERAGED ACROSS ALL WEEKDAYS AND WEEKENDS ")

```












